---
title: "Land Use and Production Time Series - BrazilAgriAnalysis"
author: "James D.A. Millington"
date: "July 2018"
output: 
  html_document: 
    keep_md: yes
    code_folding: hide
    df_print: paged
---

```{r results='hide', warning=F, message=F}

rm(list=ls())
library(tidyverse)
library(readxl)

```

Using Summary Table data for observed Land Cover (PastureB classification)
```{r warning=F, message=F}
yrs <- seq(2000, 2015, 1)

#load the region file (used to match each cell to a municipality)
region <- read.csv("SummaryTables/region.csv")

i <- 1


for(i in seq_along(yrs)){
  
  
  if(i == 1){
    lc_dat <- read_csv(paste0("SummaryTables/LCs",yrs[i],"_PastureB.csv"))   #load empirical map summary data (created using summarise_LCmaps.r)
    lc_dat <- lc_dat %>%
      mutate(year = yrs[i]) 
  }
  
  else {
    
    lc <- read_csv(paste0("SummaryTables/LCs",yrs[i],"_PastureB.csv"))   #load empirical map summary data (created using summarise_LCmaps.r)
    lc <- lc %>%
      mutate(year = yrs[i])
    
    lc_dat <- bind_rows(lc_dat, lc)
  }
  
}  

#add state ID
lc_dat <- lc_dat %>%
  mutate(state = (muniID %/% 100000)) %>%
  mutate(state = if_else(state == 17, "TO", 
      if_else(state == 29, "BA",
      if_else(state == 31, "MG",
      if_else(state == 35, "SP",
      if_else(state == 41, "PR",
      if_else(state == 42, "SC",
      if_else(state == 43, "RS", 
      if_else(state == 50, "MS",
      if_else(state == 51, "MT",
      if_else(state == 52, "GO", "NA"
      ))))))))))
    )

#LC1 = Nature
#LC2 = Other Agri
#LC3 = Agri
#LC4 = Other
#LC5 = Pasture

#add cell count columns  
lc_dat <- lc_dat %>%
      mutate(Nature = round(LC1 * NonNAs,0)) %>%
      mutate(OAgri = round(LC2 * NonNAs,0)) %>%
      mutate(Agri = round(LC3 * NonNAs,0)) %>%
      mutate(Other = round(LC4 * NonNAs,0)) %>%
      mutate(Pasture = round(LC5 * NonNAs,0))


lc_cells_long <- lc_dat %>%
  select(year:Pasture) %>%
  gather(key = LC, value = cells, -year, -state)

lc_cells_long_states <- lc_cells_long %>% 
  group_by(state, year,LC) %>%
  summarise_at(vars(matches("cells")),sum)

```


Plot observed land cover by state through time
```{r warning=F, message=F}

lc_cells_long_states %>% 
  ggplot(aes(x = year, y = cells, fill = LC)) + 
  geom_bar(stat = "identity") +
  scale_y_continuous(name = "Cells", labels = scales::comma) +
  facet_grid(.~state) +
  ggtitle("Ten States")


```


Plot observed land cover for all 10 simulation states through time
```{r warning=F, message=F}

lc_cells_long_brazil <- lc_cells_long_states %>% 
  group_by(year,LC) %>%
  summarise_at(vars(matches("cells")),sum)

lc_cells_long_brazil %>% 
  ggplot(aes(x = year, y = cells, fill = LC)) + 
  geom_bar(stat = "identity", colour="white") +
  scale_y_continuous(name = "Cells", labels = scales::comma) +
  #facet_grid(.~state) +
  ggtitle("Ten States")


```



Use production data from IBGE
```{r warning=F, message=F}
##Load Production Data
meat_prod_Astates_Data <- read_excel("Cattle_meat_production_Kg_2000_2017_all_states.xlsx", sheet = "Plan1", skip = 1)  #data for all states Astates

#dairy data are by municiaplity for all states (Amunis)
dairy_prod_Amunis_Data <- read_excel("dairy_Municipalities_Brazil.xlsx", sheet = "Tabela", skip = 1, na = c("", "-", "..."))

#maize data are by municiaplity for all states (Amunis)
maize_prod_Amunis_Data <- read_excel("maize_brazil.xlsx", sheet = "Production (tons)", skip = 1, na = c("", "-", "..."))

#maize data are by municiaplity for all states (Amunis)
soy_prod_Amunis_Data <- read_excel("soybean_brazil.xlsx", sheet = "Production (Tons)", skip = 1, na = c("", "-", "..."))


Fstate_vals <- c(17,	29,	31,	35,	41,	42,	43,	50,	51,	52)
Fstate_abbrev <- c("TO", "BA", "MG", "SP", "PR",  "SC", "RS", "MS", "MT", "GO")

Astate_codes <- meat_prod_Astates_Data %>%
  select(NM_UF_SIGLA, CD_GCUF) %>%
  rename(state = NM_UF_SIGLA, stateid = CD_GCUF) %>%
  filter(!is.na(state))    #safer way to remove text line at bottom of state column
  

##MEAT
meat_prod_Astates <- meat_prod_Astates_Data %>%
  rename(state = NM_UF_SIGLA) %>%
  select(-NM_UF, -CD_GCUF, -`2016`, -`2017`) %>%      #drop columns
  filter(!is.na(state)) %>%   #safer way to remove text line at bottom of state column
  mutate_at(vars("2000":"2015"), as.numeric) 

meat_prod_Astates_long <- meat_prod_Astates %>%
   gather(key = year, value = meat_kg, -state) %>%
   mutate_at(vars(year), as.integer) %>%
   mutate(meat_gg = meat_kg * 0.000001) %>%  #convert from kg to gg
   select(-meat_kg)

##DAIRY
dairy_prod_Amunis <- dairy_prod_Amunis_Data %>%
  rename(muniID = `IBGE CODE`) %>%
  filter(!is.na(muniID)) %>%   #safer way to remove text line in muniID
  mutate(state = substr(muniID, 1, 2)) %>%     #extract the muniID
  mutate_at(vars("2000":"2015"), as.numeric) %>%  #convert values to numeric
  select(-Municipality)   #drop unwanted columns

dairy_prod_Astates <- dairy_prod_Amunis %>%
  group_by(state) %>%
  mutate_all(funs(. * 1.03 * 1000)) %>%     #convert from thousand litres to kgs
  summarise_all(sum, na.rm=T) %>%    #summarise munis to states
  mutate(state=replace(state, 1:length(Astate_codes$stateid), Astate_codes$state)) #re-label stated ids with state abbrevs

dairy_prod_Astates_long <- dairy_prod_Astates %>%
  gather(key = year, value = dairy_kg, -state, -muniID) %>%
  mutate_at(vars(year), as.integer) %>%
  mutate(dairy_gg = dairy_kg * 0.000001) %>%  #convert from kg to gg
  select(-dairy_kg, -muniID)


#MAIZE
#has the same data strucutre (with some differences in unit conversions - could write function to cover both?) 
maize_prod_Amunis <- maize_prod_Amunis_Data %>%
  rename(muniID = `IBGE CODE`) %>%
  filter(!is.na(muniID)) %>%   #safer way to remove text line in muniID
  mutate(state = substr(muniID, 1, 2)) %>%     #extract the muniID
  mutate_at(vars("2000":"2015"), as.numeric) %>%  #convert values to numeric
  select(-Municipality)   #drop unwanted columns

maize_prod_Astates <- maize_prod_Amunis %>%
  group_by(state) %>%
  summarise_all(sum, na.rm=T) %>%    #summarise munis to states
  mutate(state=replace(state, 1:length(Astate_codes$stateid), Astate_codes$state)) #re-label stated ids with state abbrevs

maize_prod_Astates_long <- maize_prod_Astates %>%
  gather(key = year, value = maize_kg, -state, -muniID) %>%
  mutate_at(vars(year), as.integer) %>%
  mutate(maize_gg = maize_kg * 0.001) %>%  #convert from tons to gg
  select(-maize_kg, -muniID)

##SOY

soy_prod_Amunis <- soy_prod_Amunis_Data %>%
  rename(muniID = `IBGE CODE`) %>%
  filter(!is.na(muniID)) %>%   #safer way to remove text line in muniID
  mutate(state = substr(muniID, 1, 2)) %>%     #extract the muniID
  mutate_at(vars("2000":"2015"), as.numeric) %>%  #convert values to numeric
  select(-Municipality)   #drop unwanted columns

soy_prod_Astates <- soy_prod_Amunis %>%
  group_by(state) %>%
  summarise_all(sum, na.rm=T) %>%    #summarise munis to states
  mutate(state=replace(state, 1:length(Astate_codes$stateid), Astate_codes$state)) #re-label stated ids with state abbrevs

soy_prod_Astates_long <- soy_prod_Astates %>%
  gather(key = year, value = soy_kg, -state, -muniID) %>%
  mutate_at(vars(year), as.integer) %>%
  mutate(soy_gg = soy_kg * 0.001) %>%  #convert from tons to gg
  select(-soy_kg, -muniID)

prod_state_year <- left_join(meat_prod_Astates_long, dairy_prod_Astates_long, by = c("year", "state"))

prod_state_year <- left_join(prod_state_year, maize_prod_Astates_long, by = c("year", "state"))

prod_state_year <- left_join(prod_state_year, soy_prod_Astates_long, by = c("year", "state"))

#add focal states indicator
prod_state_year <- prod_state_year %>%
  mutate(simulated = state %in% Fstate_abbrev) 

psy_long <- prod_state_year %>%
  gather(key = commodity, value = gg, -state, -year, -simulated)

psimy_long <- psy_long %>%
  group_by(simulated, year, commodity) %>%
  summarise(sumsim = sum(gg, na.rm=T))

```

```{r warning=F, message=F}

psimy_long_sim <- psimy_long %>% 
  filter(simulated == "TRUE") 

psimy_long_sim %>% 
  ggplot(aes(x = year, y = sumsim, fill = commodity)) + 
  geom_bar(stat = "identity", colour="white") +
  scale_y_continuous(name = "Production (gg)", labels = scales::comma) +
  #facet_grid(.~state) +
  ggtitle("Production Data")

psimy_long_sim %>% 
  ggplot(aes(x = year, y = sumsim, colour=commodity)) + 
  geom_line(size = 1) +
  scale_y_continuous(name = "Production (gg)", labels = scales::comma) +
  #facet_grid(.~state) +
  ggtitle("Production Data")

```


For CRAFTY we assume the following values are the maximum possible production yields (gg per 25sq km):
- Soy = 20	
- Maize = 30
- Milk = 2.5
- Meat = 0.275

Using these values we can check what the observed production quantities would imply for the number of land cover cells.

```{r warning=F, message=F}

#THESE ARE PERFECT VALUES! Assumes service production = 1 
psimy_long_sim  <- psimy_long_sim %>%
  mutate(cells = if_else(commodity == "dairy_gg", sumsim / 2.5, 
      if_else(commodity == "maize_gg", sumsim / 30,
      if_else(commodity == "meat_gg", sumsim / 0.275, sumsim / 20)
      )))


psimy_long_sim %>% 
  ggplot(aes(x = year, y = cells, fill = commodity)) + 
  geom_bar(stat = "identity", colour="white") +
  scale_y_continuous(name = "Production (cells)", labels = scales::comma) +
  #facet_grid(.~state) +
  ggtitle("Production Data")


psimy_long_sim %>% 
  ggplot(aes(x = year, y = cells, colour = commodity)) + 
  geom_line(size = 1) +
  scale_y_continuous(name = "Production (cells)", labels = scales::comma) +
  #facet_grid(.~state) +
  ggtitle("Production Data")


```

From the observed production data we see increasing required cells, from 20,000 to 40,000

Now let's see what we got in the observed land cover data (we'll aggrgeate the production data to 'Agri' and Pasture' to match the LC data)
```{r warning=F, message=F}

psimy_wide_sim  <- psimy_long_sim %>%
  select(-sumsim) %>%
  spread(commodity, cells) %>%
  mutate(Agri = maize_gg + soy_gg, Pasture = dairy_gg + meat_gg)

plcsim_long_sim <- psimy_wide_sim %>%
  select(year, Agri, Pasture) %>%
  gather(key = LC, value = cells, -year, -simulated) 


plcsim_long_sim %>% 
  ggplot(aes(x = year, y = cells, fill = LC)) + 
  geom_bar(stat = "identity", colour="white") +
  scale_y_continuous(name = "Cells", labels = scales::comma) +
  #facet_grid(.~state) +
  ggtitle("Production Data")

c <- lc_cells_long_brazil %>% 
  filter(LC == "Agri" | LC == "Pasture") %>%
  ggplot(aes(x = year, y = cells, fill = LC)) + 
  geom_bar(stat = "identity", colour="white") +
  scale_y_continuous(name = "Cells", labels = scales::comma) +
  #facet_grid(.~state) +
  ggtitle("Land Cover Data")
print(c)
```

Note that proportions look about right but there are many more cells in the LC data than predicted by the conversion from production data. 

Let's check the proportions (AP ratio is ```Agri / Pasture```)

Production Data
```{r warning=F, message=F}

psimy_wide_sim <- psimy_wide_sim %>%  
  mutate(APratio = round(Agri / Pasture,3))

psimy_wide_sim

```

Land Cover Data
```{r warning=F, message=F}

lc_cells_wide_brazil <- lc_cells_long_brazil %>%
  spread(LC, cells) %>%
  mutate(APratio = round(Agri / Pasture,3))

lc_cells_wide_brazil

```

So proportions seem to be not too bad. But absolute values are way off. This is likely because observed land was not producing 'perfect' yields. So, let's see what assuming production conversion is not perfect does.

Let's assume cells on average produce 50% of perfect production:

```{r warning=F, message=F}

plcsim_long_sim %>%
  mutate(cells = cells * 2) %>%
  ggplot(aes(x = year, y = cells, fill = LC)) + 
  geom_bar(stat = "identity", colour="white") +
  scale_y_continuous(name = "Cells", labels = scales::comma) +
  #facet_grid(.~state) +
  ggtitle("Production Data (50% conversion)")

```
Now we get similar values for 2013 onwards, but not in earlier years. The increase through time (compared to LC data) likely indicates yields have improved through time... 

Let's plot the relationship between production (cells) and land cover (cells): 
```{r warning=F, message=F}

joined <- left_join(plcsim_long_sim, lc_cells_long_brazil, by = c("year", "LC"))

joined <- joined %>%
  mutate(prod_cells = round(cells.x,0)) %>% 
  select(-cells.x) %>%
  rename(lc_cells = cells.y)

joined %>% ggplot(aes(x = prod_cells, y = lc_cells, colour=year, shape=LC)) + 
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "black") +
  ggtitle("Perfect Conversion")

```

We see that for maximum (perfect) production, neither Agri nor Pasture correlate well. 

```{r}

p <- joined %>%
  mutate(prod_cells50 = prod_cells / 0.5) %>%
  ggplot(aes(x = prod_cells50, y = lc_cells, colour=year, shape=LC)) + 
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "black") +
  ggtitle("Imperfect Conversion (50%)")
print(p)

```

When we plot the 50% conversion (above) we see Agri matches up much better (at any point in time).

Pasture has a poor relationship because production is relatively poor in earlier years. So the influence of improving yield seems important for Pasture but less so Agri... 

```{r}

p <- joined %>%
  mutate(prod_cells40 = prod_cells / 0.4) %>%
  filter(LC == "Agri") %>%
  ggplot(aes(x = prod_cells40, y = lc_cells, colour=year, shape=LC)) + 
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "black") +
  ggtitle("Imperfect Conversion (40%)")
print(p)

```

When we examine 40% conversion we see that Agri is well aligned; this implies we should aim for mean Agri service provsion of 0.4.... although this will need to take into account Soy, Maize and Double Cropper services. How to do this? Need to check current mean service values...

Questions:
- is it right that yields of agri have not improved through time?; 
Well, yes we can see this from figures above... gg of both soy and maize have increased as well as overall Agri land cover area (cells). But pasture land cover (cells) has increased only a little relative to production (gg) 

So why have pasture yields increased through time? Or has it been grassland that has been converted to pasture? This wouldn't be picked up by our land cover maps data.  Could we check this issue by comparing land cover data from maps and pasture area from IBGE? We have planted/harvested data for soy and maize (which we could compare to LC maps) but not for dairy and meat...


2019-02-21
Picking this up again, to try to understand how to convert above production info into CRAFTY demand. 

First, let's calculate the 'production efficiency' through time. i.e. divide expected cells (given 'perfect' yield) by observed cells. This is how close to 'perfect' yield we see through time. 

```{r}

joined <- joined %>%
  mutate(prod_efficiency = prod_cells / lc_cells) 

p <- joined %>%
  ggplot(aes(x = year, y = prod_efficiency, colour=LC)) +
  geom_line()

print(p)
```

Now divide (perfect) production cells by this efficiency to get the number of cells we would expect to see. 

```{r}

joined <- joined %>%
  mutate(ef_cells = prod_cells / prod_efficiency)

p <- joined %>%
  ggplot(aes(x = year, y = ef_cells, colour=LC)) +
  geom_line()

print(p)
   

```

This still underestimates Agri and Pasture that we see in the real world. 

But why would we expect Supply = Demand? Producers can only respond to demand so quickly and supply may never equal demand... really demand is shown through price data. Should we be looking at Price data to estimate demand? 